EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-06-01 01_intro.tex
v1 + few corrections Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\section{Introduction and Motivation}\label{SecIntro}
Quantum impurity models play a central role in the study of strongly correlated electron systems, capturing the coupling of a small number of localized degrees of freedom with an extended non-interacting environment \cite{Nozieres1980JP,Hewson1993}. 
Originally introduced to describe diluted magnetic impurities in metallic hosts, single-impurity models have led to the full understanding of the Kondo problem \cite{Anderson1961PR,Kondo1964POTP,Schrieffer1966PR} progressively becoming powerful tools for exploring a wide range of quantum many-body phenomena \cite{Wilson1975RMP,Georges1996RMP,Kotliar2004PT,Kotliar2006RMP}. More recently, their importance has extended beyond the realm of condensed matter, influencing the understanding of other areas such as quantum information \cite{Su2013MPLB,Walsh2019PRL,Walsh2020PQ,Walsh2021PNAS,Bellomia2024PRB} or ultra-cold atomic systems \cite{Dao2007PRL,Amaricci2014PRA,Del-Re2018PRA,Walsh2019PRB,Tusi2022NP}.

At the heart of quantum impurity models is the interaction between a localized ``impurity''  and a surrounding bath of itinerant particles, usually represented as a conduction band of electrons \cite{Hewson1993}. In this context, the definition of the impurity can be taken as rather general, in order to contain orbital and spin local degrees of freedom as well as a finite number of lattice sites. 
This setup captures many essential aspects of strong correlation including screening \cite{Roekeghem2014PRL,Roekeghem2014EL,Werner2016JOPCM,Tomczak2017TEPJST}, decoherence and entanglement in a computationally tractable way \cite{Walsh2021PNAS,Bellomia2024PRB}.
Most notably, quantum impurity problems lie at the core of many quantum embedding methods, like density-matrix embedding theory (DMET) \cite{Knizia2012PRL}, Gutzwiller approximation (GA) \cite{Lanata2015PRX,Mejuto-Zaera2023PRB},
Variational Cluster Approximations \cite{Potthoff2003TEPJBCMACS,Senechal2008,Potthoff2011ACP,Nuss2011,Dionne2023SPC,Dionne2023SPCa}
or Dynamical Mean-Field Theory (DMFT) \cite{Georges1996RMP, Kotliar2004PT,Kotliar2006RMP}.
Within these approaches the complex physics of the lattice systems get reduced to an effective self-consistent quantum impurity problem.
By accounting for local quantum fluctuations, DMFT successfully describes the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds or unconventional superconductors.

Accurately and efficiently solving a quantum impurity problem and computing the Green's functions, which are central in the DMFT framework, remains though a
critical challenge, particularly for multi-orbital systems, systems
at low temperatures or endowed with lower symmetry.
%
To this end, over many years many advanced numerical techniques have been developed which incorporate cutting-edge strategies to reduce computational loads and leverage state-of-the-art algorithms \cite{Bauer2011JOSMTAE,Parcollet2015CPC}. 
These range from continuous-time quantum Monte Carlo 
methods \cite{Gull2011RMP,Rubtsov2005PRB,Haule2007PRB,Seth2016CPC,Wallerberger2019CPC},
to numerical renormalization group approaches \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Debertolis2021PRB} 
and density-matrix renormalization group \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Nunez-Fernandez2025A} or Configuration Interaction \cite{Zgid2012PRB,Lu2014PRB,Go2017PRB,Bi2019CPC,Mejuto-Zaera2019PRB} solvers. 
In this context, the Exact Diagonalization (ED) approach plays a significant 
and distinctive role \cite{Caffarel1994PRL,Dolfen2006,Perroni2007PRB,Capone2007PRB,Weber2012PRB,Lu2017TEPJST,Amaricci2022CPC}.

More in general, several open-source ED software packages have recently been made available  within the condensed matter community to address quantum many-body problems with an emphasis on efficiency and flexibility. Examples include \texttt{EDLib} \cite{Iskakov2017} and the newer 
\texttt{XDiag} \cite{Wietek2025}, which focus on generic fermionic systems; 
\texttt{Pomerol} \cite{Antipov2015} designed for fermion-boson models and interfaced with TRIQS \cite{Parcollet2015CPC}; \texttt{H$\Phi$} \cite{Kawamura2017CPC,Ido2024CPC} 
for fermions coupled to spins; and general-purpose spin-model solvers such as 
\texttt{QuSpin} \cite{Weinberg2017SP,Weinberg2019SP}, all contributing to the 
rich landscape of ED-based techniques.

%EDIPACK
In this work we present the renewed and improved \NAME{}: a flexible, high-performance numerical library designed to address these challenges, providing an exact diagonalization (ED) solver for generic Quantum Impurity Problems. 
The library builds on the foundations of
its predecessor \cite{Amaricci2022CPC}, featuring massively parallel Lanczos-based algorithms \cite{Lanczos1950JRNBSB,Lehoucq1998,Maschhoff1996} while including several significant extensions and improvements. The library now supports, within a unified framework, both zero and finite temperature calculations \cite{Amaricci2022CPC,Capone2007PRB} for a wide set of impurity models, making it capable of addressing a large set of problems, from spin-orbit coupling effects in quantum materials to multi-orbital superconductivity. It also provides direct
access to numerically exact dynamical correlation functions across the
entire frequency complex plane, thus enabling calculations of spectral functions, local susceptibilities, and other response properties.

Additionally, \NAME offers access to the Fock space of
the impurity system and thus, for instance, the possibility to evaluate the impurity reduced density matrix (iRDM). This capability enables quantum information-inspired analyses providing insights into the entanglement properties, subsystem purity, and quantum correlations of interacting electron systems \cite{Walsh2021PNAS,Bellomia2024PRB}.
In the quantum embedding context, these features are increasingly recognized as essential to understand the emergent behavior of correlated matter, including quantum criticality and superconductivity.
% MC this last point is intrinsically DMFT (or embedding) and it has no connection with impurity physics. Right: I added an opening which should clarify the sentence.

This new \NAME version also includes support for systems with non-equivalent impurities. This is a critical feature in DMFT or, more generally, in quantum embedding methods, to study hetero-structures, super-lattices, and disordered systems, where the local environment varies significantly across different sites.
% This capability makes it 
% possible to explore a wide range of inhomogeneous systems, including
% oxide interfaces or disordered systems, where
% translational symmetry is broken.
%MC? Repetition.. removed

%\paragraph{Interoperability}\\
With the growing importance of  quantum impurity models in the study of correlated materials, standardization and interoperability between solution software suites has
become a critical necessity.
\NAME has been expressly designed following this guiding principle. 
While the library itself is implemented using modern Fortran constructs, it also provides an extensive set of C-bindings that enable integration with other programming languages,
be they compiled (C, C++), JIT-compiled (Julia) or interpreted (Python).  
Especially notable in this sense is the Python API, EDIpack2py, which beyond providing a 
high-level interface for defining and solving impurity problems, also allows integration with established scientific libraries, such as NumPy and SciPy, facilitating data analysis and
rapid prototyping.

A key development feature in \NAME, unlocked by the introduction of the Python API, 
is the possibility to directly interface to both TRIQS \cite{Parcollet2015CPC} (Toolbox for Research on Interacting Quantum Systems) and
W2Dynamics \cite{Wallerberger2019CPC} software suites, which are two of the most important and widely used computational framework in the field.
\NAME slots in as an alternative solver to the native Quantum Monte Carlo 
implementations \cite{Gull2011RMP,Seth2016CPC}, extending the parameter (and especially temperature) range reachable by 
the simulation while minimizing the coding requirements for the end user.
At the same time, this integration provides \NAME users with access to the
extensive data handling and analysis tools in the TRIQS ecosystem. 
Overall, interoperability extends the cumulative reach and trustworthiness of all these platforms,
enhancing reproducibility and cross-validation of numerical results, with the ultimate goal of achieving standardization for quantum impurity solvers along the guiding principles of initiatives such as FAIRmat (see \href{https://www.fairmat-nfdi.eu/fairmat/}{fairmat-nfdi.eu/fairmat}).

%\paragraph{closure}
In this paper, we present the mathematical framework, implementation
details, and key features of \NAME, along with several illustrative
applications. We aim to demonstrate how this flexible,
high-performance solver can be used to tackle a wide range of
challenging problems. 
The rest of this work is organized as follow. In \secu{SecInstall} we 
briefly illustrate the overall structure of \NAME together with its dependencies, configuration
and installation procedures. In \secu{SecEDIpack} we present the quantum impurity
problem and review the software implementation in detail. In \secu{SecInterop} we thoroughly discuss the C-binding interface which is at the heart of the 
interoperability capabilities of \NAME. In this section we also discuss 4 interface layers: i) 
the Python API, ii) the TRIQS interface, iii) the W2Dynamics interface and iv) an experimental, yet operative, Julia API.
Finally, in \secu{SecExamples} we discuss the functionalities and the capabilities of \NAME and  all the discussed extensions with a series of illustrative examples.

>>>>

\section{Introduction and Motivation}\label{SecIntro}
Quantum impurity models play a central role in the study of strongly correlated electron systems, capturing the coupling of a small number of localized degrees of freedom to an extended non-interacting environment \cite{Nozieres1980JP,Hewson1993}. 
Originally introduced to describe diluted magnetic impurities in metallic hosts, single-impurity models have led to the full understanding of the Kondo problem \cite{Anderson1961PR,Kondo1964POTP,Schrieffer1966PR} have progressively become powerful tools for exploring a wide range of quantum many-body phenomena \cite{Wilson1975RMP,Georges1996RMP,Kotliar2004PT,Kotliar2006RMP}. More recently, their importance has expanded beyond the realm of condensed matter, influencing the understanding of other areas such as quantum information \cite{Su2013MPLB,Walsh2019PRL,Walsh2020PQ,Walsh2021PNAS,Bellomia2024PRB} or ultra-cold atomic systems \cite{Dao2007PRL,Amaricci2014PRA,Del-Re2018PRA,Walsh2019PRB,Tusi2022NP}.

At the heart of quantum impurity models is the interaction between a localized ``impurity''  and a surrounding bath of itinerant particles, usually represented as a conduction band of electrons \cite{Hewson1993}. In this context, the  impurity  can be defined more generally, so as to include orbital and spin local degrees of freedom as well as a finite number of lattice sites. 
This setup captures many essential aspects of strong correlation, including screening \cite{Roekeghem2014PRL,Roekeghem2014EL,Werner2016JOPCM,Tomczak2017TEPJST}, decoherence and entanglement in a computationally tractable way \cite{Walsh2021PNAS,Bellomia2024PRB}.
Most notably, quantum impurity problems lie at the core of many quantum embedding methods, like density-matrix embedding theory (DMET) \cite{Knizia2012PRL}, Gutzwiller approximation (GA) \cite{Lanata2015PRX,Mejuto-Zaera2023PRB},
Variational Cluster Approximations \cite{Potthoff2003TEPJBCMACS,Senechal2008,Potthoff2011ACP,Nuss2011,Dionne2023SPC,Dionne2023SPCa}
or Dynamical Mean-Field Theory (DMFT) \cite{Georges1996RMP, Kotliar2004PT,Kotliar2006RMP}.
Within these approaches the complex physics of the lattice systems gets reduced to an effective self-consistent quantum impurity problem.
By accounting for local quantum fluctuations, DMFT successfully describes the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds and unconventional superconductors.

Accurately and efficiently solving a quantum impurity problem and computing the Green's functions, which are central in the DMFT framework, nonetheless remains a critical challenge, particularly for multi-orbital systems, systems
at low temperatures or endowed with lower symmetry.
%
To this end, over the years 
many advanced numerical techniques have been developed, incorporating cutting-edge strategies to reduce computational loads and leverage state-of-the-art algorithms \cite{Bauer2011JOSMTAE,Parcollet2015CPC}. 
These range from continuous-time quantum Monte Carlo 
methods \cite{Gull2011RMP,Rubtsov2005PRB,Haule2007PRB,Seth2016CPC,Wallerberger2019CPC},
to numerical renormalization group approaches \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Debertolis2021PRB} 
and density-matrix renormalization group \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Nunez-Fernandez2025A} or Configuration Interaction \cite{Zgid2012PRB,Lu2014PRB,Go2017PRB,Bi2019CPC,Mejuto-Zaera2019PRB} solvers. 
In this context, the Exact Diagonalization (ED) approach plays a significant 
and distinctive role \cite{Caffarel1994PRL,Dolfen2006,Perroni2007PRB,Capone2007PRB,Weber2012PRB,Lu2017TEPJST,Amaricci2022CPC}.

More generally, several open-source ED software packages have recently been made available  within the condensed matter community to address quantum many-body problems with an emphasis on efficiency and flexibility. Examples include \texttt{EDLib} \cite{Iskakov2017} and the newer 
\texttt{XDiag} \cite{Wietek2025}, which focus on generic fermionic systems; 
\texttt{Pomerol} \cite{Antipov2015} designed for fermion-boson models and interfaced with TRIQS \cite{Parcollet2015CPC}; \texttt{H$\Phi$} \cite{Kawamura2017CPC,Ido2024CPC} 
for fermions coupled to spins; and general-purpose spin-model solvers such as 
\texttt{QuSpin} \cite{Weinberg2017SP,Weinberg2019SP}, all contributing to the 
rich landscape of ED-based techniques.

%EDIPACK
In this work we present the improved and renewed \NAME{}: a flexible, high-performance numerical library designed to address these challenges, providing an exact diagonalization (ED) solver for generic Quantum Impurity Problems. 
The library builds on the foundations of
its predecessor \cite{Amaricci2022CPC}, featuring massively parallel Lanczos-based algorithms \cite{Lanczos1950JRNBSB,Lehoucq1998,Maschhoff1996} while including several significant extensions and improvements. The library now supports, within a unified framework, both zero and finite temperature calculations \cite{Amaricci2022CPC,Capone2007PRB} for a wide set of impurity models, making it suitable for addressing a broad class of problems, from spin-orbit coupling in quantum materials to multi-orbital superconductivity. It also provides direct
access to numerically exact dynamical correlation functions across the
entire frequency complex plane, thus enabling calculations of spectral functions, local susceptibilities, and other response properties.

Additionally, \NAME offers access to the Fock space of
the impurity system and thus, for instance, the possibility to evaluate the impurity reduced density matrix (iRDM), enabling quantum information-inspired analyses providing insights into the entanglement properties, subsystem purity, and quantum correlations of interacting electron systems \cite{Walsh2021PNAS,Bellomia2024PRB}.
In the quantum embedding context, these features are increasingly recognized as essential to understand the emergent behavior of correlated matter, including quantum criticality and superconductivity.
% MC this last point is intrinsically DMFT (or embedding) and it has no connection with impurity physics. Right: I added an opening which should clarify the sentence.

This new \NAME version also includes support for systems with inequivalent impurities. This is a critical feature in DMFT or, more generally, in quantum embedding methods, to study hetero-structures, super-lattices, and disordered systems, where the local environment varies significantly across different sites.
% This capability makes it 
% possible to explore a wide range of inhomogeneous systems, including
% oxide interfaces or disordered systems, where
% translational symmetry is broken.
%MC? Repetition.. removed

%\paragraph{Interoperability}\\
With the growing importance of quantum impurity models in the study of correlated materials, standardization and interoperability across software suites has
become a critical necessity.
\NAME has been expressly designed following this guiding principle. 
While the library itself is implemented using modern Fortran constructs, it also provides an extensive set of C-bindings that enable integration with other programming languages,
be they compiled (C, C++), JIT-compiled (Julia) or interpreted (Python).  
Especially notable in this sense is the Python API, EDIpack2py, which beyond providing a 
high-level interface for defining and solving impurity problems, it also allows integration with established scientific libraries, such as NumPy and SciPy, facilitating data analysis and
rapid prototyping.

A key development feature in \NAME, unlocked by the introduction of the Python API, 
is the possibility to directly interface to both TRIQS \cite{Parcollet2015CPC} (Toolbox for Research on Interacting Quantum Systems) and
W2Dynamics \cite{Wallerberger2019CPC} software suites, which are two of the most important and widely used computational frameworks in the field.
\NAME slots in as an alternative solver to the native Quantum Monte Carlo 
implementations \cite{Gull2011RMP,Seth2016CPC}, extending the parameter (and especially temperature) range reachable by 
the simulation while minimizing the coding requirements for the end user.
At the same time, this integration provides \NAME users with access to the
extensive data handling and analysis tools in the TRIQS ecosystem. 
Overall, interoperability enhances the collective reach and reliability of these platforms, 
improving reproducibility and cross-validation of numerical results, with the ultimate goal of achieving standardization for quantum impurity solvers along the guiding principles of initiatives such as FAIRmat (see \href{https://www.fairmat-nfdi.eu/fairmat/}{fairmat-nfdi.eu/fairmat}).

%\paragraph{closure}
In this paper, we present the mathematical framework, implementation
details, and key features of \NAME, along with several illustrative
applications. We aim to demonstrate how this flexible,
high-performance solver can be used to tackle a wide range of
challenging problems. 
The rest of this work is organized as follows. In \secu{SecInstall} we 
briefly illustrate the overall structure of \NAME together with its dependencies, configuration
and installation procedures. In \secu{SecEDIpack} we present the quantum impurity
problem and review the software implementation in detail. In \secu{SecInterop} we thoroughly discuss the C-binding interface which is at the heart of the 
interoperability capabilities of \NAME. In this section we also discuss 4 interface layers: i) 
the Python API, ii) the TRIQS interface, iii) the W2Dynamics interface and iv) an experimental, yet operative, Julia API.
Finally, in \secu{SecExamples} we discuss the functionalities and the capabilities of \NAME and  all the discussed extensions with a series of illustrative examples.

### 2025-06-01 02_install.tex
v1 complete Full revision. Syntax check (but more checks to do). References fully added. Color codes set. Freezing this.
v1 + few corrections Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\documentclass[edipack_sp.tex]{subfiles}
\begin{document}
\fixme{AA: reviewed. Misses possible References}

\section{Installation}\label{SecInstall}
The configuration and installation of \NAME is handled by CMake, which ensures
multi-platform compatibility and dependencies resolution.  
The software builds into two distinct libraries.
The main one is a static Fortran library named {\tt libedipack.a} which, alongside the compiled
modules, wraps the \NAME software.
A second dynamic library, {\tt libedipack\_cbindings.so}, together with the associated header
file {\tt edipack\_cbindings.h}, enables interoperability with other programming languages.

\subsection{Structure}\label{sSecInstallStructure}
\NAME is a modular library organized into two main components. At its 
core, constituting the primary computational engine, 
is the  Exact Diagonalization solver for the quantum single-impurity Hamiltonian.
Building on this, the library includes 
\NAME{2ineq}, an extension for handling multiple inequivalent impurity problems at once. Finally, a Fortran-C/C++ interface is provided for 
seamless integration with external software and the development of additional APIs.

\begin{itemize}
\item{\bf EDIpack.}
  This module forms the foundation of the library, implementing the 
  Lanczos-based solver for general quantum impurity problems. It 
  supports systems with a wide range of symmetries leveraging different quantum numbers conservation, multi-orbital models, and even coupling to local 
  phonons. The \NAME solver is structured hierarchically and modularly, 
  with different sections of the library communicating through a shared 
  memory layer. The top-level module, {\tt EDIPACK}, provides access 
  to the core Fortran API, exposing key procedures for initialization, 
  execution, and finalization, while abstracting the underlying data 
  structures. A detailed overview of this part of the library is 
  provided in \secu{SecEDIpack}.
  \begin{itemize}
    \item[$\hookrightarrow$]
    {\it EDIpack2ineq.}
  This extension, developed using suitable Fortran interfaces,
  enables the treatment of multiple, independent quantum 
impurities. It is particularly useful in DMFT applications involving 
unit cells with inequivalent atomic sites or systems with broken 
translational symmetry, such as heterostructures, large supercells, 
or disordered materials. This module provides flexible memory 
management and supports the simultaneous solution of multiple impurity 
problems, as discussed in \secu{sSecIneq}.
\end{itemize}

@@

\subsection{Dependencies}\label{sSecInstallDependencies}
\NAME directly depends on two external libraries.
\begin{enumerate}
\item {\bf SciFortran}: an open-source Fortran library that provides support
for mathematical and scientific software development, available at \href{https://github.com/SciFortran/SciFortran}{github.com/SciFortran}.
\item {\bf MPI}: a distributed memory parallel communication layer with support for modern Fortran compiler.
\end{enumerate}

{\bf SciFortran} provides a solid development platform enabling access to
many algorithms and functions, including standard linear algebra
operations and high-performance Lanczos-based algorithms. This
greatly reduces code clutter and development time.
The use of distributed memory parallel environment 
is required to access scalable parallel diagonalization algorithms,
which speed up calculations for large systems. Nonetheless, through careful design \NAME operates serially even without initialization of the MPI framework.

\subsection{Build and Install}\label{sSecInstallBuildInstall}
\subsubsection{Compilation from source}
The software can be installed from source as follows. The code package can
be retrieved directly from its GitHub repository, for instance using
\begin{lstlisting}[style=mybash,numbers=none]
git clone https://github.com/edipack/EDIpack 
\end{lstlisting}
Then, assuming being in the software root directory ({\tt \$cd EDIpack}), a conventional out-of-source building can be performed using two alternative compilation backends.

@@

\noindent
The CMake configuration can be further tuned using the following variables:
\begin{center}
\begin{tabular}{ l|l|l } 
 \hline
  {\bf Option}               & {\bf Scope} & {\bf Value (default in {\color{xkcdEmerald}green})}\\
  \hline
  -D{\bf PREFIX}          & Install directory  & {\color{xkcdEmerald} $\sim$/opt/EDIpack/TAG/PLAT/BRANCH} \\
  -D{\bf USE\_MPI}       & MPI support  &  True/{\color{xkcdEmerald}False}\\
  -D{\bf WITH\_INEQ}   & multi-impurities support & {\color{xkcdEmerald}True}/{False}\\
  -D{\bf VERBOSE}      & Verbose CMake output & {\color{xkcdEmerald}True}/{False}\\ 
  -D{\bf BUILD\_TYPE} & Compilation flags & {\color{xkcdEmerald}RELEASE}/TESTING/DEBUG \\
 \hline
\end{tabular}
\end{center}
%
The default target builds and installs both the main library and the C-bindings.
Separate build targets for either components are available. A recap message is printed at the end of the CMake configuration step to guide the user through the compilation, installation and OS loading process.

\subsubsection{Anaconda}
As an alternative, we provide for both Linux and macOS systems
installation through Anaconda packages into a virtual
environment containing Python ($\geq3.10$).
%
The Conda package installation procedure for a virtual environment called {\tt myenv} reads
\begin{lstlisting}[style=mybash,numbers=none]
conda create -n myenv
conda activate myenv
conda install -c conda-forge -c edipack edipack
\end{lstlisting}
and installs a bundle of SciFortran and \NAME libraries together with
specific {\tt pkg-config} configuration files, which can be used to
retrieve compilation and linking flags. In order to compile Fortran/C++ 
drivers, the {\tt compilers} conda package will need to be installed.

\subsection{Environment Variables}\label{sSecInstallOSloading}
In order to avoid possible conflicts or the requirement for administrative
privileges, the results of the building step are installed by default in user's {\tt HOME}
directory, specified by the CMake variable {\tt PREFIX}.
As a consequence, the environment variables holding the library and include paths will need to be updated by the user.
We offer different ways to perform this action:
\begin{enumerate}
\item  A CMake-generated configuration file for an environment module
  which allows to load and unload the library at any time. This is
  preferred solution for HPC systems. 
\item A CMake-generated bash script to be sourced (once or
  permanently) in any shell session to add \NAME library to the
  default environment.
\item A CMake generated {\tt pkg-config} file to be added in
  the {\tt PKG\_CONFIG\_PATH} itself.  
\end{enumerate}
An automatically generated recap message with all instructions is
generated at the end of the installation procedure.

>>>>

\documentclass[edipack_sp.tex]{subfiles}
\begin{document}

\section{Installation}\label{SecInstall}
The configuration and installation of \NAME is handled by CMake, which ensures
cross-platform compatibility and dependency resolution.  
The software builds into two distinct libraries.
The main one is a static Fortran library named {\tt libedipack.a}, which alongside the compiled
modules, wraps the \NAME software.
A second dynamic library, {\tt libedipack\_cbindings.so}, together with the associated header
file {\tt edipack\_cbindings.h}, enables interoperability with other programming languages.

\subsection{Structure}\label{sSecInstallStructure}
\NAME is a modular library organized into two main components. 
At its core lies the ED solver for the quantum
single-impurity Hamiltonian, constituting the primary computational engine.
Building on this, the library includes \NAME{2ineq}, an extension for handling multiple inequivalent impurity problems at once. Finally, a Fortran-C/C++ interface is provided for seamless integration with external software and the development of additional APIs.

\begin{itemize}
\item{\bf EDIpack.}
  This module forms the foundation of the library, implementing a 
  Lanczos-based solver for general quantum impurity problems. It 
  supports systems with a wide range of symmetries leveraging conservation of different quantum numbers, multi-orbital models, and even coupling to local 
  phonons. The \NAME solver is structured hierarchically and modularly, 
  with different sections of the library communicating through a shared 
  memory layer. The top-level module, {\tt EDIPACK}, provides access 
  to the core Fortran API, exposing key procedures for initialization, 
  execution, and finalization, while abstracting the underlying data 
  structures. A detailed overview of this part of the library is   provided in \secu{SecEDIpack}.
  \begin{itemize}
    \item[$\hookrightarrow$]
    {\it EDIpack2ineq.}
  This extension, developed using suitable Fortran interfaces,
  enables the treatment of multiple, independent quantum 
impurities. It is particularly useful in DMFT applications involving 
unit cells with inequivalent atomic sites or systems with broken 
translational symmetry, such as heterostructures, large supercells, 
or disordered materials. This module provides flexible memory 
management and supports the simultaneous solution of multiple impurity 
problems, as discussed in \secu{sSecIneq}.
\end{itemize}

@@

\subsection{Dependencies}\label{sSecInstallDependencies}
\NAME directly depends on two external libraries.
\begin{enumerate}
\item {\bf SciFortran}: an open-source Fortran library that provides support
for mathematical and scientific software development, available at \href{https://github.com/SciFortran/SciFortran}{github.com/SciFortran}.
\item {\bf MPI}: a distributed memory parallel communication layer with support for modern Fortran compilers.
\end{enumerate}

{\bf SciFortran} provides a solid development platform enabling access to
many algorithms and functions, including standard linear algebra
operations and high-performance Lanczos-based algorithms. This
greatly reduces code clutter and development time.
The use of distributed memory parallel environment 
is required to access scalable parallel diagonalization algorithms,
which speed up calculations for large systems. Nonetheless, thanks to careful design, \NAME can operate serially even without initializing the MPI framework.

\subsection{Build and Install}\label{sSecInstallBuildInstall}
\subsubsection{Compilation from source}
The software can be installed from source as follows. The code package can
be retrieved directly from its GitHub repository, for instance using
\begin{lstlisting}[style=mybash,numbers=none]
git clone https://github.com/edipack/EDIpack 
\end{lstlisting}
Then, assuming to be in the software root directory ({\tt \$cd EDIpack}), a conventional out-of-source building can be performed using two alternative compilation backends.

@@

\noindent
The CMake configuration can be further tuned using the following variables:
\begin{center}
\begin{tabular}{ l|l|l } 
 \hline
  {\bf Option}               & {\bf Scope} & {\bf Value (default in {\color{xkcdEmerald}green})}\\
  \hline
  -D{\bf PREFIX}          & Install directory  & {\color{xkcdEmerald} $\sim$/opt/EDIpack/TAG/PLAT/BRANCH} \\
  -D{\bf USE\_MPI}       & MPI support  &  True/{\color{xkcdEmerald}False}\\
  -D{\bf WITH\_INEQ}   & multi-impurities support & {\color{xkcdEmerald}True}/{False}\\
  -D{\bf VERBOSE}      & Verbose CMake output & {\color{xkcdEmerald}True}/{False}\\ 
  -D{\bf BUILD\_TYPE} & Compilation flags & {\color{xkcdEmerald}RELEASE}/TESTING/DEBUG \\
 \hline
\end{tabular}
\end{center}
%
The default target builds and installs both the main library and the C-bindings.
Separate build targets for each component are available. A recap message is printed at the end of the CMake configuration step to guide the user through the compilation, installation and OS loading process.

\subsubsection{Anaconda}
Installation is also available through Anaconda packages for Linux and macOS systems, within a virtual environment containing Python
($\geq$ 3.10).
%
The Conda package installation procedure for a virtual environment called {\tt myenv} reads
\begin{lstlisting}[style=mybash,numbers=none]
conda create -n myenv
conda activate myenv
conda install -c conda-forge -c edipack edipack
\end{lstlisting}
and installs a bundle of SciFortran and \NAME libraries together with
specific {\tt pkg-config} configuration files, which can be used to
retrieve compilation and linking flags. In order to compile Fortran/C++ 
drivers, the {\tt compilers} conda package will need to be installed.

\subsection{Environment Variables}\label{sSecInstallOSloading}
In order to avoid possible conflicts or the requirement for administrative
privileges, the results of the building step are installed by default in user's {\tt HOME}
directory, specified by the CMake variable {\tt PREFIX}.
As a consequence, the environment variables holding the library and include paths will need to be updated by the user.
We offer different ways to perform this action:
\begin{enumerate}
\item  A CMake-generated configuration file for an environment module which allows users to load and unload the library at any time. This is
  preferred solution for HPC systems. 
\item A CMake-generated bash script to be sourced (once or
  permanently) in any shell session to add \NAME library to the
  default environment.
\item A CMake generated {\tt pkg-config} file to be added in
  the {\tt PKG\_CONFIG\_PATH} itself.  
\end{enumerate}
A recap message with all instructions is automatically generated at the end of the installation procedure.
